---
title: "HW3_Porter_Erica"
author: "Erica Porter"
date: "9/13/2017"
output: pdf_document
---
```{r setup, include=FALSE, message=F, warning=F}
options(stringsAsFactors = FALSE)
library(tidyverse)
library(beeswarm)
library(stargazer)
library(data.table)
library(lubridate)
library(lintr)
library(gridExtra)
library(plotpc)
library(xtable)
library(numDeriv)
library(data.table)
library(ggplot2)
library(vioplot)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=T, message=F, warning=F)
```

#Problem 3

Takeaways from the style guides include:
\begin{itemize}
    \item Consistency is important; it is best to develop a sufficiently neat programming style and use a similar format           throughout multiple files/projects.
    \item Spacing is essential for writing and reading code; be wary of using tabs.
    \item Identifiers and variable names should not have excessive special characters and should follow neat, standard             naming conventions
    \item Try to neatly indent code/functions/processes with different types of commands (e.g. when using tidyr, each              "verb" should be on its own line).
\end{itemize}
I will try to improve my own code by:
\begin{itemize}
    \item Weave my code and text a more effectively for R Markdown documents requiring code, text, and explanations.
    \item Format, label tables and printed results better (e.g. explore grid functions, stargazer, graph packages).
    \item Keeping track of variable and data frame names better.
    \item Including effective spacing and code chunks to improve the appearance and progression of .Rmd files, rather              than solely creating a neat PDF.
\end{itemize}

#Problem 4

```{r Problem4_lintr_practice, echo=F, eval=F}
## Problem 4 ##
## Get stylistic comments about homework 2 ##

lint(filename = ("02_data_munging_summarizing_R_git/HW2_Porter_Erica.Rmd")
```

Some suggestions for stylistic improvements included:
\begin{itemize}
    \item Inserting spaces after commas and around operators
    \item Format functions better (e.g. opening and closing curly braces on their own line)
    \item Limit lines to <80 characters
    \item Use <- for assignment rather than =
    \item Consistently use double-quotes rather than single-quotes
    \item Avoid using absolute paths if possible
    \item Variable and function names should be lowercase
    \item Avoid using absolute paths
\end{itemize}

#Problem 5
```{r Problem5_function, echo=FALSE, eval=TRUE}
## Problem 5 ##
## Write a function to determine summary statistics for device measurements ##

dev_data <- readRDS("./HW3_data.rds")

# Initialize empty vectors to receive each of the stats for the observations #
mean1 <- c()
mean2 <- c()
sd1 <- c()
sd2 <- c()
corr <- c()

frame <- as.data.frame(matrix(NA,nrow=13,ncol=6))

dev <- function(data,observers) {
  temp <- c(mean(subset(data, Observer == observers)$dev1), mean(subset(data, Observer == observers)$dev2), sd(subset(data, Observer == observers)$dev1), sd(subset(data, Observer == observers)$dev2), cor(subset(data, Observer == observers)$dev1, subset(data, Observer == observers)$dev2))
  return(c(observers,temp))
}

# Loop through the data to evaluate mean, sd, and correlation for each of 13 observers #
for(i in 1:13) {
    frame[i,] <- dev(data=dev_data,observers=i)
}

summary_stats <- cbind(mean1, mean2, sd1, sd2, corr)
summary_stats <- as.data.frame(summary_stats)
```

```{r Problem5_table, echo = F, eval= T}
## Problem 5 ##
## Paste the summary statistics into a data frame to print ##
knitr::kable(summary_stats, caption="Compare dev means")
```

```{r Problem5_graphs, echo = FALSE, eval = TRUE}
## Problem 5 ##
## Create boxplots to compare dev1 and dev2 means ##
    par(mfrow = c(1,2))
    boxplot(mean1, data = summary_stats)
    boxplot(mean2, data = summary_stats)
    vioplot(summary_stats$sd1, col = "gray")
    vioplot(summary_stats$sd2, col = "gray")
```

#Problem 6
The Blood Pressure data from Wu and Humada needs to be reformatted/tidied because the measurements for blood pressure span six different columns, the devices measurements are not grouped together, and doctor measurements are not all grouped together.  I will use the \texttt{gather}, \texttt{separate}, and \texttt{mutate} commands in \texttt{tidyr} to create a single column for day (the original data has an extraneous column for day) and columns for measuring entity, associated measure number, and the measurement value.  Below I have printed the first 5 rows of the tidy data set and a summary table describing the data.  See the Appendix for full R code.

```{r Problem6_tidy_practice, echo = F, eval = TRUE}
## Problem 6 ##
## Tidy the Blood Pressure data set ##
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BloodPressure.dat"
blood_pressure <- read.table(url, header = F, skip = 1, fill = T, stringsAsFactors = F)
blood_pressure_tidy <- blood_pressure[-1, -5]
    colnames(blood_pressure_tidy)<-c("Day", "Dev_1", "Dev_2", "Dev_3", "Doc_1", "Doc_2", "Doc_3")
    blood_pressure_tidy <- blood_pressure_tidy %>%  
        gather(measure_num, value, Dev_1:Doc_3) %>%  
        separate(measure_num,into=c("method", "replicate"), sep="_") %>%
        mutate(value = as.numeric(value))
    
# Print first 5 observations and a summary table
knitr::kable(head(blood_pressure_tidy, n = 5), caption = "First 5 observations for Blood Pressure")
knitr::kable(summary(blood_pressure_tidy), caption = "Summary of Blood Pressure data")
```

#Problem 7
```{r Problem7_newton, echo = F, eval = T}
## Problem 7 ##

# This is my first attempt at a function for Newton's method #
# This generates a solution without displaying intermediate iterations #
# Begin with a starting value for x
# Obtain an estimate x - f(x)/f'(x)
# Repeat while the tolerance conditions are satisfied

fun <- function(x) {3^x - sin(x) + cos(5*x)}
der <- function(x) {log(3)*3^x - 5*sin(5*x) - cos(x)}
newton <- function(fun, der, a, t = 0.01) {
  b <- a - fun(a)/der(a)
while ((abs(a-b) > t) & (abs(a-b)/(abs(a)+abs(b))) > t) {
    c <- a - fun(a)/der(a)
    b <- a
    a <- c
result <- a
return(result)
}}
```

```{r Problem7_newton2, echo = F, eval = T}
## Problem 7 ##
# This is my second attempt at Newton's method with iterations #

fun <- function(x) {3^x - sin(x) + cos(5*x)}
der <- function(x) {log(3)*3^x - 5*sin(5*x) - cos(x)}

newton2 <- function(f, a, b, t, n = 1000) {
x_0 <- a
  k <- n
    for (i in 1:n) {
    deriv <- der(x_0) 
    c <- x_0 - (fun(x_0) / deriv)
    k[i] <- c
    if (abs(c - x_0) < t) {
      estimate <- tail(k, n = 1)
      to_print <- list("solution" = estimate, "number iterations" = k)
      return(to_print)
      
    }
    x_0 <- c
  }
}

newton2(fun, -10, 10, 0.0001, n = 100)
```

```{r Problem8_car_data, echo = FALSE, eval = TRUE, include = T}
## Problem 8 ##
## Merge and analyze car data ##

# Read in data using fread since files are large
Car_Geb<- fread(input = "Open_Data_RDW__Gebreken.csv", header = T) #dat2
Car_Gec <- fread(input = "Open_Data_RDW__Geconstateerde_Gebreken.csv", header=T) #dat3
Car_Person <- fread(input ="Personenauto_basisdata.csv", header=T) #dat1
    
# Merge the 3 data sets be common columns
plates <- merge(Car_Person,Car_Gec,by="Kenteken")
defects <- merge(plates,Car_Geb,by="Gebrek identificatie")

# Select only columns of interest, rename columns in English
defects_small <- defects[, c(1,2,4,5,16,24)]
colnames(defects_small) <- c("Defect Code", "License", "Make", "Model", "Inspection Date", "Defect Description")

# Check for NA values in the data set we will be working with
check_na <- sum(is.na(defects_small$`Defect Code`))

# Subset all rows from year 2017 and count unique Makes and Models
defects17 <- defects_small[grep("2017",defects_small$`Inspection Date`),]
unique_make <- length(unique(defects17$Make))
unique_model <- length(unique(defects17$Model))
knitr::kable(as.data.frame(cbind(unique_make,unique_model)), caption = "Number of Unique Makes and Models")

# Find top 5 defect codes
common_defects <- as.data.frame(defects17 %>% count(`Defect Code`,sort = TRUE))
top_defects <- common_defects$`Defect Code`[1:5]

# Find most common make for each of above defects
defect_make1 <- (as.data.frame(subset(defects17, `Defect Code` == top_defects[1]) %>% count(Make, sort = TRUE)))[1,1]
defect_make2 <- (as.data.frame(subset(defects17, `Defect Code` == top_defects[2]) %>% count(Make, sort = TRUE)))[1,1]
defect_make3 <- (as.data.frame(subset(defects17, `Defect Code` == top_defects[3]) %>% count(Make, sort = TRUE)))[1,1]
defect_make4 <- (as.data.frame(subset(defects17, `Defect Code` == top_defects[4]) %>% count(Make, sort = TRUE)))[1,1]
defect_make5 <- (as.data.frame(subset(defects17, `Defect Code` == top_defects[5]) %>% count(Make, sort = TRUE)))[1,1]
top_defect_makes <- c(defect_make1,defect_make2,defect_make3,defect_make4,defect_make5)

# Find most common model for each of above makes
def1 <- subset(defects17, `Defect Code` == top_defects[1] & Make == top_defect_makes[1])
def2 <- subset(defects17, `Defect Code` == top_defects[2] & Make == top_defect_makes[2])
def3 <- subset(defects17, `Defect Code` == top_defects[3] & Make == top_defect_makes[3])
def4 <- subset(defects17, `Defect Code` == top_defects[4] & Make == top_defect_makes[4])
def5 <- subset(defects17, `Defect Code` == top_defects[5] & Make == top_defect_makes[5])

defect_model1 <- (as.data.frame(def1 %>% count(Model,sort = TRUE)))[1,1]
defect_model2 <- (as.data.frame(def2 %>% count(Model,sort = TRUE)))[1,1]
defect_model3 <- (as.data.frame(def3 %>% count(Model,sort = TRUE)))[1,1]
defect_model4 <- (as.data.frame(def4 %>% count(Model,sort = TRUE)))[1,1]
defect_model5 <- (as.data.frame(def5 %>% count(Model,sort = TRUE)))[1,1]
top_defect_models <- c(defect_model1,defect_model2,defect_model3,defect_model4,defect_model5)

# Table of common defect codes, their makes, and models
defect_table <- as.data.frame(cbind(top_defects,top_defect_makes,top_defect_models))
knitr::kable(defect_table, caption = "Top Defects, their makes, and models")

# Check for relationship b/w number of defects and make, then by model
carmake <- as.data.frame(defects17 %>% count(Make, sort = TRUE))
carmake$Make <- as.numeric(factor(carmake$Make))
colnames(carmake) <- c("Make","n")
carmodel <- as.data.frame(defects17 %>% count(Model, sort = TRUE))
carmodel$Model <- as.numeric(factor(carmodel$Model))
```

```{r linearfit, echo = FALSE, eval = TRUE, include = TRUE, results='asis'}
## Problem 8 ##
fit = lm(n~Make, data = carmake)
stargazer(fit,title = "Description of linear fit by Make",header = F,no.space=T,single.row=T)
stargazer(anova(fit),title = "ANOVA of linear fit by Make",header = F,no.space=T,single.row=T)


fit1 = lm(n ~ Model, data = carmodel)
stargazer(fit1,title = "Description of linear fit by Model",header = F,no.space=T,single.row=T)
stargazer(anova(fit1),title = "ANOVA of linear fit by Model",header = F,no.space=T,single.row=T)
```
There appears to be a relationship between car make and number of defects.

This workflow could certainly be improved; I had trouble finding the most common makes and models for each defects, so I ended up creating many separate data sets and using \texttt{subset} and \texttt{count} multiple times.  This does not seem very efficient to me and I would have liked to use the \texttt{subset} and \texttt{count} functions to make the table of common values in one step (or at least fewer).
\newpage

#Appendix: R code
```{r Appendix, ref.label = "Problem4_lintr_practice", echo=TRUE, eval=F, tidy=TRUE, include=T}
```
\vspace{1cm}
```{r Appendix, ref.label = "Problem5_function", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```

```{r Appendix, ref.label = "Problem5_table", echo = TRUE, eval = F, tidy = TRUE, include = T}
```

```{r Appendix, ref.label = "Problem5_graphs", echo = TRUE, eval = F, tidy = TRUE, include = T}
```

\vspace{1cm}
```{r Appendix, ref.label = "Problem6_tidy_practice", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```

```{r Appendix, echo = TRUE, eval = FALSE, tidy = TRUE, inlcude = TRUE}

## Another method I tried for tidying Blood Pressure Data that I wanted to keep for reference ##
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BloodPressure.dat"
blood_pressure <- read.table(url, header = F, skip = 1, fill = T, stringsAsFactors = F)
blood_pressure_tidy <- blood_pressure[-1,]
colnames(blood_pressure_tidy) <- c("Day_1", "Dev_1", "Dev_2", "Dev_3", "Day_2", "Doc_1", "Doc_2", "Doc_3")
    devices <- blood_pressure_tidy[, c("Dev_1", "Dev_2", "Dev_3")]
    doctors <- blood_pressure_tidy[, c("Doc_1", "Doc_2", "Doc_3")]
    devices <- devices %>%
        gather(Device, device_read, Dev_1:Dev_3) %>%
        mutate(Device = gsub("Dev_","",Device))
    doctors <- doctors %>%
        gather(Doctor, doc_read, Doc_1:Doc_3) %>%
        mutate(Doctor = gsub("Doc_","",Doctor))
    Day <- rep(1:15,3)
blood_pressure_tidy <- cbind(Day, devices, doctors)
blood_presure_tidy <- as.numeric(c(blood_pressure_tidy$device_read,blood_pressure_tidy$doc_read))
```

\vspace{1cm}
```{r Appendix, ref.label = "Problem7_newton", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```

```{r Appendix, ref.label = "Problem7_newton2", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```

\vspace{1cm}
```{r Appendix, ref.label = "Problem8_car_data", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```

```{r Appendix, ref.label = "linearfit", echo = TRUE, eval = FALSE, tidy = TRUE, include = TRUE}
```
