---
title: "R Notebook"
output: html_notebook
---

#Problem 2
```{r Problem2_sumsquares, echo = F, eval = T}

# Sum of squares for loop
set.seed(12345)
y <- seq(from = 1, to = 100, length.out = 1e+08) + rnorm(1e+08)

sum_sq = 0
diff_sq <- vector()
meany<-mean(y)
system.time({for (i in 1:length(y)) {
  diff_sq[i] <- (y[i] - meany)^2
  sum_sq = sum_sq + diff_sq[i]
}})


# Sum of squares with matrices
SSvec <- vector()

system.time(SSvec <- sum((y-meany)^2))

```

```{r Problem3, echo = F, eval = T}
set.seed(1256)
theta <- as.matrix(c(1, 2), nrow = 2) 
X <- cbind(1, rep(1:10, 10))
h <- X %*% theta + rnorm(100, 0, 0.2)

first_theta <- matrix(0, nrow = 2)
alpha = 0.01
tol = 10e-05
m <- length(h)
# i = 0
# h_0 <- first_theta + theta * x
# h is y
# h0(x) = Θ0 + Θ1x
# (abs(a-b) > t) & (abs(a-b)/(abs(a)+abs(b))) > t)
for (i in 1:m) {
while (abs(theta[2] - first_theta[2]) > tol && abs(theta[1] - first_theta[1]) > tol) {
    
    first_theta[1] <- theta[1] - (alpha * (1/m)) * sum(first_theta[1] + first_theta[2] * X[i,2] - h[i])
    first_theta[2] <- theta[2] - (alpha * (1/m)) * sum((theta[1] + theta[2] * X[i,2] - h[i]) * X[i,2])
    
}}

# Compare tp
lm(h ~ 0 + X)
```
The result from the loop above is very close to lm(h ~ 0 + X)

#Problem 4
Inverting a matrix and then applying matrix multiplication takes a significant amount of time, so John Cook suggests that rather than using the $\texttt{solve}$ function to find the inverse of $X'X$ and then multiply by $X'y$, use the $\texttt{solve}$ function to directly solve for $\hat{\beta}$.  So, instead of $solve(X'X)$ multiplied by $X'y$, combine the $solve$ and multiplication into one step with $solve(X'X, X'y)$.  I'm not sure what is happening other than eliminating a matrix multiplication, but it seems that possible that R can use the factors/solutions already calculated from each x to somehow "backsubstitute" to solve for additional solutions.  I imagine it being a little like solving a system of individual linear equations.

#Problem 5
```{r Problem5, echo = F, eval = T}
set.seed(12456)
G <- matrix(sample(c(0, 0.5, 1), size = 16000, replace = T), ncol = 10)
R <- cor(G) # R: 10 * 10 correlation matrix of G
C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
id <- sample(1:16000, size = 932, replace = F)
q <- sample(c(0, 0.5, 1), size = 15068, replace = T) # vector of length 15068 
A <- C[id, -id] # matrix of dimension 932 * 15068
B <- C[-id, -id] # matrix of dimension 15068 * 15068
p <- runif(932, 0, 1)
r <- runif(15068, 0, 1)
C <- NULL #save some memory space

sizeA <- object.size(A)
sizeB <- object.size(B)

# y = p + A*B*(q − r)
```